### 2024-02-16 documentation/thesis/thesis_abstract.tex
.

For this use case, the MagneticReadoutProcessing library has been
created in this work. It implements all important aspects of
acquisition, storage and analysis, and each intermediate step can be
customised by the user without having to create everything from scratch,
thus encouraging exchange between different user groups.

@@

To evaluate the software framework and hardware, two digital sensors
were characterised and tested under the same conditions for noise,
temperature dependence and linearity. The result is that both sensors
can only be used under certain conditions, as the noise intensity is
higher than the specified 50 \(\mu\)T and the value ranges fluctuate too
much as a result. It is recommended to consider analogue sensors or NMR
probes in the future.

>>>>

For this use case, the python based MagneticReadoutProcessing library
has been created in this work. It implements all important aspects of
acquisition, storage and analysis, and each intermediate step can be
customised by the user without having to create everything from scratch,
thus encouraging exchange between different user groups.

@@

To evaluate the software framework and hardware, two digital sensors
were characterised and tested under the same conditions for noise,
temperature dependence and linearity. The result is that both sensors
can only be used under certain conditions, as the noise intensity is
higher than the specified 50 \(\mu\)T and the value ranges fluctuate too
much as a result. It is recommended to consider the implementation of
additional digital sensors as well as analogue sensors or NMR probes in
the future.

### 2024-02-16 documentation/thesis/thesis_document.tex
.

The focus of this thesis is to improve low-frequency \gls{mri}
technology by examining the usability of magnetic field sensors for
characterising permanent magnets used in these systems. The variability
in the strength of permanent magnets leads to significant difficulties
in constructing an \gls{mri} with the necessary precision for homogenous
field generation.

@@

The Halbach magnet incorporated in this system boasts a 27cm diameter, a
\emph{B0} field strength of 50.4mT, and an impressive homogeneity of
2400\gls{ppm} over a 20cm diameter using smaller magnets
(\(12 x 12 x 12 mm^3\)). This exceptional homogeneity enables the
utilization of coil-based gradients for spatial encoding, significantly
enhancing the flexibility of image acquisition.

@@

Two sensors have been meticulously chosen for inclusion in the study.
The ultimate objective is to assess whether these selected sensors align
with the stringent criteria of achieving an accuracy level of
1000\gls{ppm}. Furthermore, the study seeks to validate whether the
measuring range of these sensors appropriately corresponds to the
required field strength, ensuring their suitability for the intended
application.

@@

The use of analog sensors is purposefully avoided, though they are
suitable for more precise measurements and extended measuring ranges.
They are excluded because they require more carefully designed circuits
and more complicated energy management.

@@

\begin{lstlisting}[language={C++}, caption={CustomSensor-Class for adding new sensor hardware support}, label=lst:CustomSensorClass]
#ifndef __CustomSensor_h__
#define __CustomSensor_h__
// register your custom sensor in implemented_sensors.h also
class CustomSensor: public baseSensor
{
public:
  CustomSensor();
  ~CustomSensor();
  // implement depending sensor communication interface
  bool begin(TwoWire& _wire_instance); // I2C
  bool begin(HardwareSerial& _serial_instance); // UART
  bool begin(Pin& _pin_instance); // ANALOG or DIGITAL PIN like onewire
  // FUNCTIONS USED BY READOUT LOGIC
  bool is_valid() override;
  String capabilities() override;
  String get_sensor_name() override;
  bool query_sensor() override;
  sensor_result get_result() override;
};
#endif
\end{lstlisting}

@@

The parameter *\_readings* should later receive the imported
measurements from the \emph{stage rawimport}
\ref{lst:pipeline_mrp_evaluation_yaml} and the optional
\emph{IP\_return\_count} parameter specifies the number of best
measurements that are returned. The return parameter is a list of
measurements containing the most similar measurements, measured by the
smallest distance between all measurements. The distance for each
measurement is determined using the centre of gravity function
\emph{CoG}, the length is then calculated from the result vector. This
value can then be used for sorting.

@@

Measuring the noise in a magnetic field sensor requires a precise
procedure and a special measurement setup. First, the magnetic field
sensor is placed in a quiet and static environment to minimize external
field interference. The temperature chamber for all noise tests is set
to \(\mu_{trev}\)=21.0\(^{\circ}\) and the sensors are placed \emph{24}
hours before the measurement run in the final measurement configuration
inside of the chamber.

The procedure starts with the acquisition of the baseline by operating
the sensor without external magnetic fields. For this purpose, a sample
size of \emph{N=10000} measured values is recorded for the baseline
measurement.

The output signal of the sensor is continuously measured and recorded.
It is important to carry out the measurement over a sufficiently long
period of time in order to record both short-term and long-term
fluctuations. For this purpose, \emph{N=2000} further measured values
are taken with a trigger and readout rate of one measurement per second.

In order to quantify the noise, the \gls{sd} of the signal is
calculated. These parameters provide information about the variation of
the signal over time and therefore about the sensor background noise.

@@

The temperature stability of the \emph{TLV493D} with a mean value of
20.68\(^{\circ}\)C and a \gls{sd} of \(\sigma_{t}\)=0.53\(^{\circ}\)C
indicates a consistent trend. This implies a constant tendency. The
close grouping of the measured values around the mean value indicates
good stability. The confidence interval is expected to be between
20.15\(^{\circ}\)C and 21.21\(^{\circ}\)C, which indicates a stable and
consistent temperature measurement. This result is more noisy compared
to the temperature stability of the \emph{MMC5603}.

Both sensors provide an offset to the measured chamber temperature
\(\mu_{trev}\).

With an additional measurement run with a different temperature setting
of 30.0\(^{\circ}\)C, the measured temperature deviations and offsets
remains constant.

The sensor internal temperature sensors of both tested sensors are
suitable to perform an ambient temperature compensation of measured
values and calibration of the sensor. This is considered in section
\emph{Temperature Sensitivity} \ref{temperature-sensitivity}.

@@

In the raw data plot, as well as its mean value \(\mu_{rv}\) of both
sensors, an offset can also be recognised, which is dependent on the
ambient conditions. For verification, a reference measurement of the
environment is carried out with the calibrated \emph{Voltcraft GM70}
Telsameter next to the sensor \gls{ic} This provides a \emph{real}
baseline value of \(\mu_{rev}\)=-21\(\mu\)T.

@@

This is achieved here by means of an additional linear axis installed
above the sensor setup. A holder for an \emph{N45 12x12x12mm} magnet is
attached to the end effector of this axis, which can thus be moved at
different distances above the respective sensor \gls{ic}. The ambient
temperature is set to \(\mu_{trev}\)=21.0\(^{\circ}\) in the measurement
runs and thus corresponds to the same conditions as in the
\emph{Background-Noise}\ref{sensor-characterisation-background-noise}
setup.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Move the linear axis upwards by the selected distance
\item
  De-energise the axis motor
\item
  Create measurement using \emph{MRPCli} with N=2000 data points per
  reading
\item
  Export reading to filesystem with current linear axis position in
  meta-data and filename
\end{enumerate}

@@

For both sensors, the deviation is less than 1\% over the entire
resolution. With the \emph{MMC5603NJ} this is on average only
\emph{0.04\%}. With the \emph{TLV493D}, however, the \gls{sd} is
\emph{3.64\%}, for which the deviations at the end in particular (with
field strengths towards zero) are decisive. The previously performed
\emph{Background-Noise} \ref{sensor-characterisation-background-noise}
characterisation shows that the linearity deviation here is due to the
sensitivity of the sensor.

@@

The same procedure is used as for the \emph{Linearity}
\ref{sensor-characterisation-linearity} measurement, except that instead
of moving the linear axis, the temperature of the temperature chamber is
systematically increased from 20\(^{\circ}\) to 50\(^{\circ}\). Between
each of these temperature changes, the system is given a waiting time of
30 minutes after reaching the target temperature.

The field of permanent magnets is very temperature-dependent and can
lose its magnetisation at higher temperatures (typically
\textgreater=80\(^{\circ}\) for non-high-quality type N magnets
\cite{magna-c}). The temperature range is selected so that it is
within a sufficient range for the application.

@@

The graph of the \emph{TLV493d} is significantly steeper with a gradient
of \emph{-5.13 \(\mu\)T / \(^{\circ}\)C}, and it is also not as linear
as the \emph{MMC5603NJ}; there are clear jumps in the gradient. However,
the total change between the temperature regions of 175\(\mu\)T is less
than that of the \emph{MMC5603NJ} with 70\(\mu\)T, if the total
measurement range of the sensors is also taken into account.

@@

The \emph{TLV493D} performs seriously in the Senosr Noise measurement
and also performs worse than specified in the data sheet (98\(\mu\)T
instead of 175\(\mu\)T), but the large measuring range, which fulfils
the required specifications from chapter \emph{Research Question}
\ref{research-question-and-approach}, must be taken into account here,
which is not met by the \emph{MMC5603NJ}.

@@

However, due to the limited measuring range of \emph{±3mT}, direct
measurement of stronger magnets is not possible. The \emph{N45
12x12x12mm} magnets used in the application typically have a field
strength of around \emph{100mT} at a distance of \emph{10mm}.

The \emph{TLV493D}, on the other hand, is able to measure these
directly, but does not achieve the required accuracy due to strong noise
and steep temperature coefficients.

In the following chapter, recommendations for action are defined, which
were derived from the analysis results.

@@

In contrast, the \emph{TLV493D} has too much noise, so that the measured
values cannot be used directly without post-processing. With software
calibration, it is probably possible to reduce the noise to below
50\(\mu\)T with the \emph{TLV493D}. As a test, several further
measurement runs were carried out, which achieved results for the sensor
noise of 71\(\mu\)T to 41\(\mu\)T when averaging \emph{100} measured
values and \emph{1000} measured values. However, further measurement
runs must be carried out to verify these results.

In general post-processing of the measurements should, include
temperature compensation with a separate temperature sensor, especially
in the case of changing measurement conditions (e.g.~movement of the
sensor in the B0 field).

>>>>

The focus of this thesis is to improve low-frequency \gls{mri}
technology by examining the usability of magnetic field sensors for
characterising permanent magnets used in these systems. The variability
in the strength of permanent magnets leads to significant difficulties
in constructing an \gls{mri} magnet with the necessary precision for
homogenous field generation.

@@

The Halbach magnet incorporated in this system boasts a \emph{27cm}
diameter, a \emph{B0} field strength of 50.4mT, and an impressive
homogeneity of 2400\gls{ppm} over a \emph{20cm} diameter using smaller
magnets (\(12 x 12 x 12 mm^3\)). This exceptional homogeneity enables
the utilization of coil-based gradients for spatial encoding,
significantly enhancing the flexibility of image acquisition.

@@

Two sensors have been meticulously chosen for inclusion in the study.
The ultimate objective is to assess whether these selected sensors align
with the stringent criteria of achieving an accuracy level of
\emph{1000}\gls{ppm}. Furthermore, the study seeks to validate whether
the measuring range of these sensors appropriately corresponds to the
required field strength, ensuring their suitability for the intended
application.

@@

The use of analogue sensors is purposefully avoided, though they are
suitable for more precise measurements and extended measuring ranges.
They are excluded because they require more carefully designed circuits
and more complicated energy management.

@@

\begin{lstlisting}[language={C++}, caption={CustomSensor-Class for adding new sensor hardware support}, label=lst:CustomSensorClass]
#ifndef __CustomSensor_h__
#define __CustomSensor_h__
// register custom sensor in implemented_sensors.h also
class CustomSensor: public baseSensor
{
public:
  CustomSensor();
  ~CustomSensor();
  // implement depending sensor communication interface
  bool begin(TwoWire& _wire_instance); // I2C
  bool begin(HardwareSerial& _serial_instance); // UART
  bool begin(Pin& _pin_instance); // ANALOG or DIGITAL PIN like onewire
  // FUNCTIONS USED BY READOUT LOGIC
  bool is_valid() override;
  String capabilities() override;
  String get_sensor_name() override;
  bool query_sensor() override;
  sensor_result get_result() override;
};
#endif
\end{lstlisting}

@@

The parameter \_readings should later receive the imported measurements
from the \emph{stage rawimport} \ref{lst:pipeline_mrp_evaluation_yaml}
and the optional \emph{IP\_return\_count} parameter specifies the number
of best measurements that are returned. The return parameter is a list
of measurements containing the most similar measurements, measured by
the smallest distance between all measurements. The distance for each
measurement is determined using the centre of gravity function
\emph{CoG}, the length is then calculated from the result vector. This
value can then be used for sorting.

@@

Measuring the noise in a magnetic field sensor requires a precise
procedure and a special measurement setup. First, the magnetic field
sensor is placed in a quiet and static environment to minimize external
field interference. The temperature chamber for all noise tests is set
to \emph{\(\mu_{trev}\)=21.0\(^{\circ}\)} and the sensors are placed
\emph{24} hours before the measurement run in the final measurement
configuration inside of the chamber.

The procedure starts with the acquisition of the baseline by operating
the sensor without external magnetic fields. For this purpose, a sample
size of \emph{N=10000} measured values is recorded for the baseline
measurement. The output signal of the sensor is continuously measured
and recorded. It is important to carry out the measurement over a
sufficiently long period of time in order to record both short-term and
long-term fluctuations. For this purpose, \emph{N=2000} further measured
values are taken with a trigger and readout rate of one measurement per
second. In order to quantify the noise, the \gls{sd} of the signal is
calculated. These parameters provide information about the variation of
the signal over time and therefore about the sensor background noise.

@@

The temperature stability of the \emph{TLV493D} with a mean value of
\emph{20.68\(^{\circ}\)C} and a \gls{sd} of
\emph{\(\sigma_{t}\)=0.53\(^{\circ}\)C} indicates a consistent trend.
This implies a constant tendency. The close grouping of the measured
values around the mean value indicates good stability. The confidence
interval is expected to be between \emph{20.15\(^{\circ}\)C} and
\emph{21.21\(^{\circ}\)C}, which indicates a stable and consistent
temperature measurement. This result is more noisy compared to the
temperature stability of the \emph{MMC5603}. Both sensors provide an
offset to the measured chamber temperature \(\mu_{trev}\).

With an additional measurement run with a different temperature setting
of \emph{30.0\(^{\circ}\)C}, the measured temperature deviations and
offsets remains constant.

The sensor internal temperature sensors of both tested sensors are
suitable to perform an ambient temperature compensation of measured
values and calibration of the sensor. This is considered in section
\emph{Temperature Sensitivity}
\ref{sensor-characterisation-temperature-sensitivity}.

@@

The noise of a sensor describes unwanted, random fluctuations or signals
in the measured data. These are clearly recognisable in the figure ** in
the raw data plot. The baseline determined for each sensor is shown by
the red ** line. Here it can be seen that each sensor has a different
mean value or baseline offset in the same environment.

The noise of a sensor describes unwanted, random fluctuations or signals
in the measured data. These are clearly recognisable in the figure
\ref{Sensor_noise_evaluation_results_for_TLV493D_and_MMC5603NJ_with_N=2000_samples_and_no_averaging.png}
in the raw data plot. The baseline determined for each sensor is shown
by the red \(\mu_{rv}\) {[}\(\mu\)T{]} line. Here can be seen that each
sensor has a different mean value or baseline offset in the same
environment. For verification, a reference measurement of the
environment is carried out with the calibrated \emph{Voltcraft GM70}
Telsameter next to the sensor \gls{ic}. This provides a reference
baseline value of \emph{\(\mu_{rev}\)=-21.0\(\mu\)T}.

The noise of the sensor is drawn around this using the blue line. For
the \emph{TLV493D} the \gls{sd} \emph{\(\sigma_{rv}\)=172.0\(^{\mu}\)T},
which is twice the value given by the datasheet as the noise value.
Under the same conditions, the \emph{MMC5603NJ} undercuts the value
specified by the manufacturer. The \gls{sd} approaches
\emph{\(\sigma_{rv}\)=0.20\(^{\mu}\)T} and is negligible, especially
when additional averaging is used.

As the baseline was determined first for each sensor, it can be seen
that the Another recognisable feature is Further measurement series were
also recorded with inserted \emph{N45 12x12x12mm} magnets, so that these
are saturated halfway through the measurement range. This allows the
noise measurement values to be compared with a blank measurement. These
values correspond, so it can be concluded that the noise is within the
same range across the measuring range.

@@

This is achieved here by means of an additional linear axis installed
above the sensor setup. A holder for an \emph{N45 12x12x12mm} magnet is
attached to the end effector of this axis, which can thus be moved at
different distances above the respective sensor \gls{ic}. The ambient
temperature is set to \emph{\(\mu_{trev}\)=21.0\(^{\circ}\)} in the
measurement runs and thus corresponds to the same conditions as in the
\emph{Background-Noise}\ref{sensor-characterisation-background-noise}
setup.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Move the linear axis upwards by the selected distance
\item
  De-energise the axis motor
\item
  Create measurement using \emph{MRPCli} with \emph{N=2000} data points
  per reading
\item
  Export reading to filesystem with current linear axis position in
  meta-data and filename
\end{enumerate}

@@

For both sensors, the deviation is less than \emph{1\%} over the entire
resolution. With the \emph{MMC5603NJ} this is on average only
\emph{0.04\%}. With the \emph{TLV493D}, however, the \gls{sd} is
\emph{3.64\%}, for which the deviations at the end in particular (with
field strengths towards zero) are decisive. The previously performed
\emph{Background-Noise} \ref{sensor-characterisation-background-noise}
characterisation shows that the linearity deviation here is due to the
sensitivity of the sensor.

@@

The same procedure is used as for the \emph{Linearity}
\ref{sensor-characterisation-linearity} measurement, except that instead
of moving the linear axis, the temperature of the temperature chamber is
systematically increased from \emph{20\(^{\circ}\)} to
\emph{50\(^{\circ}\)}. Between each of these temperature changes, the
system is given a waiting time of 30 minutes after reaching the target
temperature.

The field of permanent magnets is very temperature-dependent and can
lose its magnetisation at higher temperatures (typically
\emph{\textgreater=80\(^{\circ}\)} for non-high-quality type N magnets
\cite{magna-c}). The temperature range is selected so that it is
within a sufficient range for the application.

@@

The graph of the \emph{TLV493d} is significantly steeper with a gradient
of \emph{-5.13 \(\mu\)T / \(^{\circ}\)C}, and it is also not as linear
as the \emph{MMC5603NJ}; there are clear jumps in the gradient. However,
the total change between the temperature regions of \emph{175\(\mu\)T}
is less than that of the \emph{MMC5603NJ} with \emph{70\(\mu\)T}, if the
total measurement range of the sensors is also taken into account.

@@

The \emph{TLV493D} performs in noise measurements worse than specified
in the data sheet (\emph{98\(\mu\)T} instead of \emph{175\(\mu\)T}), but
the large measuring range, which fulfils the required specifications
from chapter \emph{Research Question}
\ref{research-question-and-approach}, must be taken into account here,
which is not met by the \emph{MMC5603NJ}.

@@

However, due to the limited measuring range of \emph{±3mT}, direct
measurement of stronger magnets is not possible using the
\emph{MMC5603NJ}. The \emph{N45 12x12x12mm} magnets used in the
application typically have a field strength of around \emph{100mT} at a
distance of \emph{10mm} which is more than the \emph{MMC5603NJ} can
measure.

The \emph{TLV493D}, on the other hand, is able to measure these ranges,
but does not achieve the required accuracy due to strong noise and steep
temperature coefficients. In the following chapter, recommendations for
action are defined, which were derived from the analysis results.

@@

In contrast, the \emph{TLV493D} has too much noise, so that the measured
values cannot be used directly without post-processing. With software
calibration, it is probably possible to reduce the noise to below
\emph{50\(\mu\)T} with the \emph{TLV493D}. As a test, several further
measurement runs were carried out, which achieved results for the sensor
noise of \emph{71}\(\mu\)T* to \emph{41\(\mu\)T} when averaging
\emph{100} measured values and \emph{1000} measured values. However,
further measurement runs must be carried out to verify these results.

In general post-processing of the measurements should, include
temperature compensation with a separate temperature sensor, especially
in the case of changing measurement conditions (e.g.~movement of the
sensor in the \emph{B0} field).
